A Graph Theory Based Approach to MDAO Problem Formulation

### 2013-03-15 graph_syntax.tex
merged

@@

A number of restrictions are places on the model node with respect to the number and 
type of edges that can be connected to it: 
\begin{enumerate}
\item A model node can only have one edge directed to or from another model node.
\item A model node can only have fixed edges directed in or out.
\item A model node must have at least one edge directed in and at least one edge directed out.
\end{enumerate}

@@

\subsection{Maximal Connectivity Graph}
To construct the maximal connectivity graph, we assume that a set of codes, global inputs, and objectives and constraints (collectively called global outputs). The codes are represented by analysis blocks $A_i=(V_{A_i},E_{A_i}), \ i=1,\ldots,m$, the global inputs are represented a set of variable nodes $I$, and the global outputs are represented by a set of variable nodes $O$. We assume that $O$, $I$, and $A_i$ are given, and that any potential connection between variables is given in the form of the free edges in the set $C_M$. 
Then we may construct the maximal connectivity graph $M=(V_M,E_M)$ as
\begin{IEEEeqnarray*}{rCl}
V_M & = & I \cup O \cup \left( \bigcup_{i = 1}^m V_{A_i} \right), \\
E_M & = & C_M \cup \left( \bigcup_{i=1}^m E_{A_i} \right),
\end{IEEEeqnarray*}
The MCG $M$ is uniquely determined by the given set of analysis blocks, the required outputs, and the given global inputs. In the cases where the set of global inputs $I$ is not known a priori, the process of obtaining the FPF will reveal the required inputs, as discussed subsequently.

\subsection{Fundamental Problem Formulation Graph}
We now define the fundamental problem formulation graph, $F=(V_F,E_F)$, as a directed graph meeting the following conditions
\begin{enumerate}
\item[(1)] $\displaystyle{V_F = I_F \cup O \cup \left( \bigcup_{i \in \mathcal A} V_{A_i} \right),\ I_F \subset I}$
\item[(2)] $\displaystyle{E_F = C_F \cup \left( \bigcup_{i \in \mathcal A} E_{A_i} \right)}$
\item[(3)] $\displaystyle{\forall v \in V_F \txt{ with } t_\txt{node}(v) = \txt{`variable,'}\quad \txt{deg}_l^-(v) < \txt{deg}^-(v) \leq \txt{deg}_u^-(v)}$
\end{enumerate}
The set $\mathcal A$ is an index set containing the indices of the analysis blocks in $F$; for the case where all of the analysis blocks are used, we would have $\mathcal A = \{1,2,\ldots,m\}$. The first requirement for $F$ is that $V_F$ be composed of all of the global outputs (the required objectives and constraints), the nodes from each analysis block in $F$, and any global input nodes that are needed, $I_F$. The second requirement suggests that the edges in $F$ comprise the edges for each analysis block and the edges between them, the global inputs, and the global outputs. The final requirement is specific to the number of edges directed into a variable. If there are no edges directed inward, the node is called a \emph{hole}, and if more connections are directed in than are allowed, the node is called a \emph{collision}; a hole is allowed if $\txt{deg}_l^-(v)=0$. The final requirement is therefore a requirement on $C_F$.

@@

\begin{enumerate}
\item The first step is to detect holes and disconnect the free edges following them. These free edges are removed because they represent variables which cannot be determined because the analysis function does not have adequate inputs. The set of variable nodes which are holes is created as
\begin{equation}
H = \{v \in V | \txt{deg}^-(v) < \txt{deg}_l^-(v) \}
\end{equation}
Then the updated set of edges is
\begin{equation}
C_{F,1} = C_{F,0} \setminus \{\ e \in E, \ e=xy | x=E^3(v) \txt{ for } v \in H\}
\end{equation}
%deg-v is now recaculated for the new C. 
This step is demonstrated by Fig.~\ref{f:holes}.
\item (())
\end{enumerate}
\begin{figure}[htb!]
	\begin{center}
	\includegraphics[width=3in]{images/analysis_block_hole}
	\end{center}
	\vspace{-20pt}
\caption{Example variable node indicating a hole.}
\label{f:holes}
\end{figure}

>>>>

Let $I$ be a nonempty set such that for each $i \in I$ there is a corresponding set $A_i$. The family $\{A_i | i \in I\}$ is an indexed family of sets with index $i$ and indexing set $I$, and we can say $\mathcal A = \{A_i | i \in I\}$\cite{smith2006}. Then the union over this family of sets is
\begin{equation}
\bigcup_{i \in I} A_i = \bigcup_{A \in \mathcal A} A = \{x | x \in A \txt{ for } A \in \mathcal A\}.
\end{equation}
Lastly, the cardinality of a set $B$ is the number of elements in $B$ and is denoted as $|B|$.

@@

A number of restrictions are placed on the model node with respect to the number and 
type of edges that can be connected to it: 
\begin{enumerate}
\item A model node can only have one edge directed to or from another model node.
\item A model node can only have fixed edges directed in or out.
\item A model node must have at least one edge directed in and at least one edge directed out.
\end{enumerate}

@@

\subsection{Maximal Connectivity Graph}
To construct the maximal connectivity graph, we assume that a set of codes, global inputs, and objectives and constraints (collectively called global outputs). The codes are represented by analysis blocks $A_i=(V_{A_i},E_{A_i}), \ i\in \{1,2,\ldots,m\}=I$, the global inputs are represented a set of variable nodes $V_\txt{in}$, and the global outputs are represented by a set of variable nodes $V_\txt{out}$. We assume that $V_\txt{in}$, $V_\txt{out}$, and $A_i$ are given, and that any potential connection between variables is given in the form of the free edges in the set $C_M$. 
Then we may construct the maximal connectivity graph $M=(V_M,E_M)$ as
\begin{IEEEeqnarray*}{rCl}
V_M & = & V_\txt{in} \cup V_\txt{out} \cup \left( \bigcup_{i \in I} V_{A_i} \right), \\
E_M & = & C_M \cup \left( \bigcup_{i \in I} E_{A_i} \right),
\end{IEEEeqnarray*}
The MCG $M$ is uniquely determined by the given set of analysis blocks, the required outputs, and the given global inputs. In the cases where the set of global inputs $I$ is not known a priori, the process of obtaining the FPF will reveal the required inputs, as discussed subsequently.

\subsection{Fundamental Problem Formulation Graph}
We now define the fundamental problem formulation graph, $F=(V_F,E_F)$, as a directed graph meeting the following conditions
\begin{enumerate}
\item[(1)] $\displaystyle{V_{\txt{in},F} = \{v \in V_\txt{in} | \txt{deg}^+(v) > 0\} }$
\item[(2)] $\displaystyle{I_F = \{i \in I | \exists v \in V_{A_i} \txt{ with } \txt{deg}^+(v)>0 \txt{ and } t_\txt{node}(u)=\txt{`model'} \txt{ for } (u,v) \in E_{A_i}\} }$
\item[(2)] $\displaystyle{V_F = V_{\txt{in},F} \cup V_\txt{out} \cup \left( \bigcup_{i \in I_F} V_{A_i} \right)}$
\item[(3)] $\displaystyle{E_F = C_F \cup \left( \bigcup_{i \in I_F} E_{A_i} \right)}$
\item[(4)] $\displaystyle{\forall v \in V_F \txt{ with } t_\txt{node}(v) = \txt{`variable,'} \  \txt{deg}_l^-(v) < \txt{deg}^-(v) \leq \txt{deg}_u^-(v)}$
\end{enumerate}
The set $I_F$ is an index set containing the indices of the analysis blocks in $F$; for the case where all of the analysis blocks are used, we would have $I_F = I = \{1,2,\ldots,m\}$. 
The set $V_{\txt{in},F}$ is the set of global inputs used by the analysis blocks in $F$. Because it is possible for some global inputs to be unused, we only require $V_{\txt{in},F} \subset V_\txt{in}$.

The first requirement creates the set $V_{\txt{in},F}$ which is the set of only the global inputs that are used.
The second requirement creates set $I_F$ which is an index set corresponding to only that analysis blocks
The first requirement for $F$ is that $V_F$ be composed of all of the global outputs (the required objectives and constraints), the nodes from each analysis block in $F$, and any global input nodes that are needed.
The second requirement suggests that the edges in $F$ comprise the edges for each analysis block and the edges between them, the global inputs, and the global outputs. 
The final requirement is specific to the number of edges directed into a variable. 
If there are no edges directed inward, the node is called a \emph{hole}, and if more connections are directed in than are allowed, the node is called a \emph{collision}; a hole is allowed if $\txt{deg}_l^-(v)=0$. 
The final requirement is therefore a requirement on $C_F$.

@@

\begin{enumerate}
\item The first step is to detect holes and disconnect the free edges following them. These free edges are removed because they represent variables which cannot be determined because the analysis function does not have adequate inputs. The set of variable nodes which are holes is created as
\begin{equation}
H = \{v \in V | t_\txt{node}(v) = \txt{`variable'} \txt{ and } \txt{deg}^-(v) < \txt{deg}_l^-(v) \},
\end{equation}
which is the set of variable nodes with few incoming edges than are allowed by the lower indegree limit.
Then the updated set of edges is created by removing the first set of free edges following the hole:
\begin{equation}
C_{F,1} = C_{F,0} \setminus \{(x,y) \in E | x=E^3(v) \txt{ for } v \in H\},
\end{equation}
%deg-v is now recaculated for the new C. 
and this step is demonstrated by Fig.~\ref{f:holes}.
\item The second step is to detect collisions and to then disconnect enough free edges so that the collisions are resolved. The set of variable nodes which containing collisions is created as
\begin{equation}
S_\txt{nodes} = \{v \in V | t_\txt{node}(v) = \txt{`variable'} \txt{ and } \txt{deg}^-(v) > \txt{deg}_u^-(v) \}.
\end{equation}
For each collision node we can construct a set contained the edges directed in. The set containing all of these sets is constructed as
\begin{equation}
S_\txt{edges} = \big \{ \{(x,y) \in E\} | y \in S_\txt{nodes} \big \}
\end{equation}

\end{enumerate}
\begin{figure}[htb!]
	\begin{center}
	\includegraphics[width=3in]{images/analysis_block_hole}
	\end{center}
	\vspace{-20pt}
\caption{Example variable node indicating a hole.}
\label{f:holes}
\end{figure}
